Classify veg

### 2020-11-11 R/classify_species.Rmd
added vertical shading to variable importance plots
to help interpret which wavelengths were important

```{r varImp_byWavelength_plot, fig.height=6, fig.width=10}
varImp_MDA_species <- ggplot(imp, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "",  # x = "Wavelength (nm)\n", # remove x axis label for the combined plot
       y = "Species\nVariable importance\n",
       title = "Variable Importance: Mean Decrease Accuracy (MDA)", # \n\nSpecies classification",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        title = element_text(size = 16))

@@

```{r varImp_Gini, fig.height=6, fig.width=10}
varImp_Gini_species <- ggplot(imp, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance \nMean Decrease Gini (MDG)\n",
       #title = "Variable Importance: Mean Decrease Gini (MDG)",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 14))

@@

```{r vertical_dashed_lines}
# plot variable importance with wavelengths of interest marked by dashed vertical lines

ggplot(imp, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) +

  # add vertical dashed lines to indicate notable wavelengths
  geom_vline(xintercept=480, linetype="dashed", color="blue", size=1, alpha=0.5) + # BLUE 
  geom_vline(xintercept=550, linetype="dashed", color="green", size=1, alpha=0.5) + # GREEN
  geom_vline(xintercept=650, linetype="dashed", color="red", size=1, alpha=0.5) + # RED
  geom_vline(xintercept=870, linetype="dashed", color="grey30", size=1, alpha=0.5) + # NIR 
  geom_vline(xintercept=1600, linetype="dashed", color="chocolate4", size=1, alpha=0.5) + # SWIR

  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDG)",
       title = "Variable Importance: Mean Decrease Gini (MDG) vs. Wavelength",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca")

```
```{r vertical_shaded_regions}
# plot variable importance with wavelengths of interest marked by dashed vertical lines

ggplot(imp, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) +

  # add vertical shaded regions to indicate notable wavelengths
  annotate(geom = "rect", xmin = 450, xmax = 510, ymin = -5, ymax = Inf, fill = "blue", alpha = 0.1) + # BLUE
  annotate(geom = "rect", xmin = 530, xmax = 590, ymin = -5, ymax = Inf, fill = "green", alpha = 0.1) + # GREEN
  annotate(geom = "rect", xmin = 640, xmax = 670, ymin = -5, ymax = Inf, fill = "red", alpha = 0.1) + # RED
  annotate(geom = "rect", xmin = 850, xmax = 880, ymin = -5, ymax = Inf, fill = "grey30", alpha = 0.1)  + # NIR 
  annotate(geom = "rect", xmin = 1570, xmax = 1650, ymin = -5, ymax = Inf, fill = "chocolate4", alpha = 0.1)  + # SWIR

  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDG)",
       title = "Variable Importance: Mean Decrease Gini (MDG) vs. Wavelength",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca")

@@

```{r varImp_MDA_genus, fig.height=6, fig.width=10}
varImp_MDA_genus <- ggplot(imp_genus, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "\nWavelength (nm)", 
       y = "Genus\nVariable importance\n",
       #title = "Genus classification",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        title = element_text(size = 16))

@@

```{r varImp_Gini_genus, fig.height=6, fig.width=10}
ggplot(imp_genus, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDG)",
       title = "GENUS classification variable importance: Mean Decrease Gini (MDG) vs. Wavelength",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
```

@@

```{r varmImp_species}
ggpubr::ggarrange(varImp_MDA_species, varImp_MDA_genus, nrow = 2)

>>>>

```{r varImp_byWavelength_plot, fig.height=6, fig.width=10}
varImp_MDA_species <- ggplot(imp, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) +

  # add vertical shaded regions to indicate notable wavelengths
  annotate(geom = "rect", xmin = 450, xmax = 510, ymin = 0, 
           ymax = max(imp$MeanDecreaseAccuracy), fill = "blue", alpha = 0.1) + # BLUE
  annotate(geom = "rect", xmin = 530, xmax = 590, ymin = 0, 
           ymax = max(imp$MeanDecreaseAccuracy), fill = "green", alpha = 0.1) + # GREEN
  annotate(geom = "rect", xmin = 640, xmax = 700, ymin = 0, 
           ymax = max(imp$MeanDecreaseAccuracy), fill = "red", alpha = 0.1) + # RED
  annotate(geom = "rect", xmin = 850, xmax = 910, ymin = 0, 
           ymax = max(imp$MeanDecreaseAccuracy), fill = "grey30", alpha = 0.1)  + # NIR 
  annotate(geom = "rect", xmin = 1580, xmax = 1640, 
           ymin = 0, ymax = max(imp$MeanDecreaseAccuracy), fill = "chocolate4", alpha = 0.1) +   # SWIR

  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "",  # x = "Wavelength (nm)\n", # remove x axis label for the combined plot
       y = "Species\nVariable importance\n",
       title = "Variable Importance: Mean Decrease Accuracy (MDA)", # \n\nSpecies classification",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        title = element_text(size = 16))

@@

```{r varImp_Gini, fig.height=6, fig.width=10, eval=FALSE, include=FALSE}
varImp_Gini_species <- ggplot(imp, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance \nMean Decrease Gini (MDG)\n",
       #title = "Variable Importance: Mean Decrease Gini (MDG)",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#43a2ca") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 14))

@@

@@

```{r varImp_MDA_genus, fig.height=6, fig.width=10}
varImp_MDA_genus <- ggplot(imp_genus, aes(x=wavelength_nm,  
                y=MeanDecreaseAccuracy)) +

  # add vertical shaded regions to indicate notable wavelengths
  annotate(geom = "rect", xmin = 450, xmax = 510, ymin = 0, 
           ymax = max(imp_genus$MeanDecreaseAccuracy), fill = "blue", alpha = 0.1) + # BLUE
  annotate(geom = "rect", xmin = 530, xmax = 590, ymin = 0, 
           ymax = max(imp_genus$MeanDecreaseAccuracy), fill = "green", alpha = 0.1) + # GREEN
  annotate(geom = "rect", xmin = 640, xmax = 700, ymin = 0, 
           ymax = max(imp_genus$MeanDecreaseAccuracy), fill = "red", alpha = 0.1) + # RED
  annotate(geom = "rect", xmin = 850, xmax = 910, ymin = 0, 
           ymax = max(imp_genus$MeanDecreaseAccuracy), fill = "grey30", alpha = 0.1)  + # NIR 
  annotate(geom = "rect", xmin = 1580, xmax = 1640, ymin = 0, 
           ymax = max(imp_genus$MeanDecreaseAccuracy), fill = "chocolate4", alpha = 0.1) +   # SWIR

  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseAccuracy), color = NA) +
  theme_minimal() + 
  labs(x = "\nWavelength (nm)", 
       y = "Genus\nVariable importance\n",
       #title = "Genus classification",
       fill = "MDA")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        title = element_text(size = 16))

@@

```{r varImp_Gini_genus, fig.height=6, fig.width=10, eval=FALSE, include=FALSE}
ggplot(imp_genus, aes(x = wavelength_nm,  
                y = MeanDecreaseGini)) + 
  # add bars where each height is equal to the variable importance metric
  geom_col(aes(fill = MeanDecreaseGini), color = NA) +
  theme_minimal() + 
  labs(x = "Wavelength (nm)", 
       y = "Variable importance (MDG)",
       title = "GENUS classification variable importance: Mean Decrease Gini (MDG) vs. Wavelength",
       fill = "MDG")  + 
  scale_fill_gradient(low = "#d9d9d9", high = "#fc8d59") + 
   theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
```

@@

```{r varmImp_species_and_genus}
ggpubr::ggarrange(varImp_MDA_species, varImp_MDA_genus, nrow = 2)
